---
title: 'Group Assignment #1'
author: "Name: __________________"
output: pdf_document
---

### Q1

- Load the internal R dataset "faithful"
    - library(datasets)
- View the dataset in the source window
    - view(faithful)
- Print only the 'waiting' variable and all associated rows to the console: *Recall*: In order to subset a dataset you can use the $ sign...for example: y$x
- *Answer*: How many observations does the 'waiting' variable contain (Hint: number of rows)? 
    - length(...)

```{r}
# ...
```


### Q2

- Load the internal R dataset "ipod"
    - library(openintro)
- View the dataset in the source window
    - view(ipod)
- *Answer*: What type of variable is 'songLength'?
    - str(...)

```{r}
# ...
```


### Q3

- Run the following code to assign two data vectors, beers and pizza
```{r}
beers <- c(4, 2, 3, 5, 9)
pizza <- c(2, 3, 2, 1, 8)
```
- Create a new vector 'bp' and assign it the values of multiplying the beers and pizza vectors and then dividing the result by 2
- *Answer*: What result of the mathematical operation above is the lowest? (The actual number calculated)

```{r}
# ...
```

### Q4 

- Run the following code to assign the data vector, z
```{r}
z <- c(4, 5, 9, 8, 2, 9)
```

- Extract the second through the fourth elements of the vector and assign it the name 'y'

```{r}
# ...
```

