---
title: "Group Assignment #4 Answer Key"
author: "Curry W. Hilton"
output: html_document
---

### Q1

A sample of 20 observations have been collected on number of beers consumed by students on home Bama baseball games. The sample mean and sample standard deviation are, $\bar{x} = 5$ and $s = 2$, respectively. The population standard deviation is *unknown* in this case. What is the 95% conﬁdence interval for the true population mean?

```{r}
n <- 20
xbar <- 5
s <- 2
se <- s/sqrt(n)
t <- qt(0.05/2, 19, lower.tail = F)
moe <- t*se
ci <- xbar + c(-1, 1)*moe
ci
```

### Q2

The overall population mean of the number of days students lift weights each week is 2.78 days with a *known* standard deviation of 2.56 days. Compute a 99% confidence interval for the true population mean when $n=30$?  

```{r}
n <- 30
xbar <- 2.78
sigma <- 2.56
se <- sigma/sqrt(n)
z <- qnorm(0.01/2, lower.tail = F)
moe <- z*se
ci <- xbar + c(-1, 1)*moe
ci
```

### Q3

The 'MLB' dataset in the openintro package refers to the salaries ("salary") of each major league baseball player in 2010 (assume population).  Follow the directions below:

a) Calculate the population parameters, $\mu$ and $\sigma$

```{r}
library(openintro)
mu <- mean(MLB$salary)
sigma <- sd(MLB$salary)
mu
sigma
```

b) Take a random sample of 100 players and assign it the name of "samp" using the following code:

```{r}
set.seed(1986)
samp <- sample(MLB$salary, 100)
```

c) Calculate the sample statistics, $\bar{x}$ and $s$ from the sample taken

```{r}
xbar <- mean(samp)
s <- sd(samp)
xbar
s
n <- length(samp)
n
```

d) Contrust the 95% confidence interval assuming the population standard deviation is *known*; Use $\sigma$

```{r}
z <- qnorm(0.05/2, lower.tail = F)
se <- sigma/sqrt(n)
moe <- z*se
ci <- xbar + c(-1, 1)*moe
ci
```


e) Construct the 95% confidence interval assuming the population standard deviation is *unknown*; Use $s$ 

```{r}
t <- qt(0.05/2, n-1, lower.tail = F)
se <- s/sqrt(n)
moe <- t*se
ci <- xbar + c(-1, 1)*moe
ci
```

f) Compare the two confidence intervals in terms of range and the critical values, $z$ and $t$.  Do both intervals "capture" the true population parameter, $\mu$?

*The 95% CI using the $z$ critical does not capture the true population parameter, $\mu$.  The 95% CI using the $t$ critical value does capture the true population paramter, $\mu$.  Although, both intervals are relatively close, the CI under the condition of an unknown standard deviation is larger than the CI under the condition of a known standard deviation.  The heavy right skewness plays a factor in this situation.*
