---
title: "Midterm Exam Review"
author: "Curry W. Hilton"
output: html_document
---

Instructions: Please select the correct answer from the choices given. Enter the “Letter” (A, B, C, or D) for multiple choice type questions and for True or False type questions, enter the “Letter” either (A or B) depending on the correct answer. This exam is due at 11:59 PM on March 10, 2016. All questions must be answered and submitted within the Midterm Exam link found on BlackBoard. No late submissions will be accepted regardless the circumstances. You are to work *independently* on this exam. You are allowed to use any resources available, notes from class, your book, the previous group assignments/problem set/labs, internet, etc. Any questions regarding wording of questions or for further clarity must be directed through me (I promise to provide a response within 1 hour from the time of the email or text). Just to reiterate, you are bound under the student conduct agreement for testing procedures and will be held accountable for any academic misconduct. Do work!

### 1.

The following data points have been collected in a sampling procedure: (6, 4, 2, 8, 5) Calculate the sample mean and sample median. For simplicity sake you can store the sample observations as the R object x with the code provided.

```{r}
x <- c(6, 4, 2, 8, 5)
```

A) 5, 2
B) 8, 5
C) 5, 5
D) 4.78, 5

-----

Use the “ballBearing” dataset in the openintro package to answer the following three questions (2, 3, 4). The variable of interest in the “ballBearing” dataset is ‘lifeSpan’.

### 2. 

The sample mean of the ‘lifeSpan’ variable is:

A) 6.89223
B) 6.84922
C) 6.72991
D) 6.82921

### 3. 

The sample median of the ‘lifeSpan’ variable is:

A) 6.72904
B) 6.89223
C) 6.72991
D) 6.82921

### 4. 

The distribution of the ‘lifeSpan’ variable is considered:

A) Skewed Right 
B) Skewed Left 
C) Symmetric 
D) Multimodal

------

Use the “cars” dataset in the openintro package to answer the following three questions (5, 6, 7, 8). The variables of interest in the “cars” dataset is ‘weight’, ‘mpgCity’, ‘price’.

### 5.

Derive the covariance between variables ‘weight’ and ‘mpgCity’. The covariance between both variables indicates what type of relationship?

A) Positive 
B) Inverse/Negative 
C) No relationship

### 6.

What is the correlation measurement between variables ‘weight’ and ‘mpgCity’?

A) 0.8769183 
B) 0.9014384 
C) -0.9014384 
D) -0.8769183

### 7.

Based on your results found in question 5 and 6, the linear strength between variables ‘weight’ and ‘mpgCity’ is considered:

A) Strong positive linear relationship 
B) Strong inverse/negative linear relationship 
C) Strong positive non-linear relationship 
D) Stong inverse/negative non-linear relationship

### 8. 

Use the following scatterplot to determine the type of relationship present between variables ‘weight’ and ‘price’.

```{r}
library(openintro)
plot(cars$weight, cars$price)
```

A) Positive 
B) Inverse/Negative 
C) No relationship

-------

### 9.

A variable representing the number of people in any given statistics course at the University of Alabama is best described as:

A) Continuous & Numeric 
B) Continuous & Discrete 
C) Categorical & Discrete 
D) Discrete & Numeric

------

Use the following chart below to answer questions 10 & 11.

```{r}
table(births$premature, births$smoke)
```

### 10.

The total number of babies born in full term when mother’s were non-smokers is:

A) 87 
B) 42 
C) 21 
D) 129

### 11.

The total number of babies born in full term is:

A) 100 
B) 129 
C) 21 
D) 150

-------

Use the following boxplot to answer question 12.

```{r}

```


